---
title: "CNA_analysis"
author: "Jakub Kubiś"
date: "24 08 2020"
output: html_document
---

###Data preparing
# Requirements 1
```{r echo=TRUE}
library(tidyverse)
```

##Data joining
#Breast and ovarian data

```{r}

breast_cna <- read.csv("DataBreast/copy_number_somatic_mutation.tsv.gz", sep = '\t', header = TRUE)
breast_cna <- breast_cna %>% 
  select(chromosome,
         chromosome_start,
         segment_mean,
         icgc_donor_id) 
  
breast_cna$cna_location <- paste0('chromosome: ',
                                  breast_cna$chromosome,' : ',
                                  breast_cna$chromosome_start)
breast_cna <- breast_cna %>% 
  select(-chromosome, -chromosome_start)


ovary_cna <- read.csv("DataOvary/copy_number_somatic_mutation.tsv.gz", sep = '\t', header = TRUE)
ovary_cna <- ovary_cna %>% 
  select(chromosome,
         chromosome_start,
         segment_mean,
         icgc_donor_id)
  
ovary_cna$cna_location <- paste0('chromosome: ',
                                  ovary_cna$chromosome,' : ',
                                  ovary_cna$chromosome_start)

ovary_cna <- ovary_cna %>% 
  select(-chromosome, -chromosome_start)

patient <- breast_cna %>% 
  inner_join(ovary_cna, by = 'cna_location') %>% 
  distinct()

patient <- patient %>% 
  select(cna_location) %>% 
  distinct()

mutation_cna <- rbind(breast_cna, ovary_cna)

mutation_cna <- mutation_cna %>% 
  filter(cna_location %in% patient$cna_location)
         

remove(breast_cna, ovary_cna, patient)


```

#Donor data

```{r}

donor_breast <- read.csv("DataBreast/donor.tsv", sep = '\t', header = TRUE)

donor_breast <- donor_breast %>% 
  select(icgc_donor_id,
         donor_vital_status, 
         disease_status_last_followup,
         donor_age_at_diagnosis,
         donor_age_at_last_followup,
         donor_sex) %>%
  filter(donor_sex == 'female')

donor_breast$cancer_type <- "Breast cancer"

donor_ovary <- read.csv("DataOvary/donor.tsv", sep = '\t', header = TRUE)

donor_ovary <- donor_ovary %>% 
  select(icgc_donor_id,
         donor_vital_status, 
         disease_status_last_followup,
         donor_age_at_diagnosis,
         donor_age_at_last_followup,
         donor_sex) %>%
  filter(donor_sex == 'female')

donor_ovary$cancer_type <- 'Ovarian cancer'
```

#Join donor - mutation

```{r}


mutation_breast_cna <- mutation_cna %>% 
  left_join(donor_breast, by = 'icgc_donor_id') 

mutation_ovary_cna <- mutation_cna %>% 
  left_join(donor_ovary, by = 'icgc_donor_id')

cna <- rbind(mutation_breast_cna, mutation_ovary_cna)  

cna <- cna %>% 
  mutate_if(is.character, as.factor) %>% 
  drop_na() 
  

remove(breast, ovary, donor_breast, donor_ovary, mutation_cna, mutation_breast_cna, mutation_ovary_cna, patient)

write.csv(cna, file = "AfterPreparingData/cna_prepared.csv")

remove(cna)


```


###Models - machine learning

#Requirements 2

```{r}
library(tidyverse)
library(rpart)
library(rpart.plot)

```

#Wide-data preparing for models

```{r}

mutation_cna_long_data <- read.csv("AfterPreparingData/cna_prepared.csv", 
                               stringsAsFactors = T, header = T, row.names = 1)



mutation_cna_long_data <- mutation_cna_long_data %>% 
  select(cna_location, donor_vital_status, segment_mean, icgc_donor_id) %>% 
  filter(donor_vital_status != "")

mutation_cna_long_data <- aggregate(mutation_cna_long_data$segment_mean, list(mutation_cna_long_data$cna_location, mutation_cna_long_data$icgc_donor_id, mutation_cna_long_data$donor_vital_status), FUN=median)

mutation_cna_long_data <- mutation_cna_long_data %>% 
  rename(cna_location = Group.1,
         icgc_donor_id = Group.2,
         donor_vital_status = Group.3,
         segment_mean = x)



mutation_cna_wide_data <- pivot_wider(mutation_cna_long_data, names_from = cna_location, values_from = segment_mean , values_fill = 0)


mutation_cna_wide_data <- mutation_cna_wide_data %>% 
  select(-icgc_donor_id) %>% 
  distinct()



remove(mutation_cna_long_data)

write.csv(mutation_cna_wide_data, file = "AfterPreparingData/mutation_cna_wide_data.csv")
```


#Pararellation

```{r}

library('doParallel')

cl <- makeCluster(4) 
registerDoParallel(cl)

```


##Hierarchic tree

```{r}

mutation_cna_wide_data <- read.csv("AfterPreparingData/mutation_cna_wide_data.csv", header = T, stringsAsFactors = T, row.names = 1)


tree_1 <-  rpart(as.matrix(mutation_cna_wide_data$donor_vital_status) ~ as.matrix(mutation_cna_wide_data[, colnames(mutation_cna_wide_data) != "donor_vital_status"]))


jpeg("Figures/cna_mean_combined.jpeg" , units="in", width=15, height=5, res=300)

rpart.plot(tree_1)



dev.off()

```

#Confusion matrix

```{r}

pred_risk_tree <- predict(object = tree_1, newdata = mutation_cna_wide_data, type = "class")

confusionMatrix(data = pred_risk_tree, reference = mutation_cna_wide_data$donor_vital_status, positive = "alive", mode = "everything")

```

##Training/Validation data dividing

```{r}


split <- createDataPartition(y = mutation_cna_wide_data$donor_vital_status, p = 0.8)


train_cancer <- mutation_cna_wide_data[split$Resample1,]
valid_cancer <- mutation_cna_wide_data[-split$Resample1,]

tree_1 <- rpart(formula = donor_vital_status ~  ., data = mutation_cna_wide_data)


pred_risk_train <- predict(object = tree_1, newdata = train_cancer, type = "class")

pred_risk_valid <- predict(object = tree_1,newdata = valid_cancer, type = "class")

```
#Training_data_confusion_matrix

```{r}

confusionMatrix(data = pred_risk_train, reference = train_cancer$donor_vital_status, positive = "alive", mode = "everything")

```

#Validation_data_confusion_matrix

```{r}

confusionMatrix(data = pred_risk_valid, reference = valid_cancer$donor_vital_status,positive = "alive", mode = "everything")
```

##Multiple Linear Regression (MLR) - checking the significance of the mutation

```{r}

mutation_cna_wide_data <- read.csv('AfterPreparingData/mutation_cna_wide_data.csv', header = T, stringsAsFactors = T , row.names = 1)


mutation_cna_wide_data <- mutation_cna_wide_data %>% 
  select('chromosome: 2 : 146877806', 'chromosome: 9 : 23363115', 'chromosome: 20 : 1561580',
         'chromosome: 8 : 39388299', 'chromosome: 8 : 115650697', 'chromosome: 6 : 53935348',
         'chromosome: 6 : 77461073', 'chromosome: 7 : 64637128', 'chromosome: 5 : 173205058',
         'chromosome: 4 : 10234214', 'chromosome: 10 : 24377674', 'chromosome: 18 : 5963763',
         'chromosome: 1 : 110242953', 'chromosome: 16 : 15048768', 'chromosome: 7 : 109453982',
         'chromosome: 12 : 9637909', 'chromosome: 3 : 192886435', 'chromosome: 3 : 195478694',
         'chromosome: 1 : 169242401', 'chromosome: 20 : 1590257', 'chromosome: 8 : 24979904',
         'chromosome: 11 : 4976865', 'donor_vital_status') 

mutation_cna_wide_data$donor_vital_status <- as.numeric(mutation_cna_wide_data$donor_vital_status)

MLR_model <- lm(as.matrix(as.numeric(mutation_cna_wide_data$donor_vital_status)) ~ as.matrix(mutation_cna_wide_data[, colnames(mutation_cna_wide_data) != "donor_vital_status"]))

summary(MLR_model)



```
#Wide-data preparing for plots

```{r}

mutation_cna_long_data <- read.csv("AfterPreparingData/cna_prepared.csv", 
                               stringsAsFactors = T, header = T, row.names = 1)



mutation_cna_long_data <- mutation_cna_long_data %>% 
  select(cna_location, donor_vital_status, segment_mean, icgc_donor_id) %>% 
  filter(donor_vital_status != "")

mutation_cna_long_data <- aggregate(mutation_cna_long_data$segment_mean, list(mutation_cna_long_data$cna_location, mutation_cna_long_data$icgc_donor_id, mutation_cna_long_data$donor_vital_status), FUN=median)

mutation_cna_long_data <- mutation_cna_long_data %>% 
  rename(cna_location = Group.1,
         icgc_donor_id = Group.2,
         donor_vital_status = Group.3,
         segment_mean = x)



mutation_cna_wide_data <- pivot_wider(mutation_cna_long_data, names_from = cna_location, values_from = segment_mean , values_fill = NA)


mutation_cna_wide_data <- mutation_cna_wide_data %>% 
  select(-icgc_donor_id) %>% 
  distinct()



remove(mutation_cna_long_data)

```

#Violin-box plots

```{r}

ggplot(data = mutation_cna_wide_data) +
  geom_violin(aes(donor_vital_status, `chromosome: 2 : 146877806`, fill = donor_vital_status), na.rm = T)+
  geom_boxplot(aes(donor_vital_status, `chromosome: 2 : 146877806`, fill = donor_vital_status), na.rm = T, width=0.1)
ggsave(filename = "Figures/alive_decease_cna_1.jpeg", dpi = 600)

ggplot(data = mutation_cna_wide_data) +
  geom_violin(aes(donor_vital_status, `chromosome: 9 : 23363115`, fill = donor_vital_status), na.rm = T)+
  geom_boxplot(aes(donor_vital_status, `chromosome: 9 : 23363115`, fill = donor_vital_status), na.rm = T, width=0.1)
ggsave(filename = "Figures/alive_decease_cna_2.jpeg", dpi = 600)

ggplot(data = mutation_cna_wide_data) +
  geom_violin(aes(donor_vital_status, `chromosome: 20 : 1561580`, fill = donor_vital_status), na.rm = T)+
  geom_boxplot(aes(donor_vital_status, `chromosome: 20 : 1561580`, fill = donor_vital_status), na.rm = T, width=0.1)
ggsave(filename = "Figures/alive_decease_cna_3.jpeg", dpi = 600)

ggplot(data = mutation_cna_wide_data) +
  geom_violin(aes(donor_vital_status, `chromosome: 5 : 173205058`, fill = donor_vital_status), na.rm = T)+
  geom_boxplot(aes(donor_vital_status, `chromosome: 5 : 173205058`, fill = donor_vital_status), na.rm = T, width=0.1)
ggsave(filename = "Figures/alive_decease_cna_4.jpeg", dpi = 600)

ggplot(data = mutation_cna_wide_data) +
  geom_violin(aes(donor_vital_status, `chromosome: 10 : 24377674`, fill = donor_vital_status), na.rm = T)+
  geom_boxplot(aes(donor_vital_status, `chromosome: 10 : 24377674`, fill = donor_vital_status), na.rm = T, width=0.1)
ggsave(filename = "Figures/alive_decease_cna_5.jpeg", dpi = 600)

ggplot(data = mutation_cna_wide_data) +
  geom_violin(aes(donor_vital_status, `chromosome: 18 : 5963763`, fill = donor_vital_status), na.rm = T)+
  geom_boxplot(aes(donor_vital_status, `chromosome: 18 : 5963763`, fill = donor_vital_status), na.rm = T, width=0.1)
ggsave(filename = "Figures/alive_decease_cna_6.jpeg", dpi = 600)

ggplot(data = mutation_cna_wide_data) +
  geom_violin(aes(donor_vital_status, `chromosome: 3 : 195478694`, fill = donor_vital_status), na.rm = T)+
  geom_boxplot(aes(donor_vital_status, `chromosome: 3 : 195478694`, fill = donor_vital_status), na.rm = T, width=0.1)
ggsave(filename = "Figures/alive_decease_cna_7.jpeg", dpi = 600)

ggplot(data = mutation_cna_wide_data) +
  geom_violin(aes(donor_vital_status, `chromosome: 20 : 1590257`, fill = donor_vital_status), na.rm = T)+
  geom_boxplot(aes(donor_vital_status, `chromosome: 20 : 1590257`, fill = donor_vital_status), na.rm = T, width=0.1)
ggsave(filename = "Figures/alive_decease_cna_8.jpeg", dpi = 600)

ggplot(data = mutation_cna_wide_data) +
  geom_violin(aes(donor_vital_status, `chromosome: 11 : 4976865`, fill = donor_vital_status), na.rm = T)+
  geom_boxplot(aes(donor_vital_status, `chromosome: 11 : 4976865`, fill = donor_vital_status), na.rm = T, width=0.1)
ggsave(filename = "Figures/alive_decease_cna_9.jpeg", dpi = 600)



```



