---
title: "Expression analysis"
author: "Jakub Kubiś"
date: "25 08 2020"
output: html_document
---

###Analysis exrpession of genes in ovarian/breast cancer

#Requirements 1

```{r}
library(tidyverse)
Sys.setenv(LANG = "en")
set.seed(123)
```

###Load and preparing data for breast cancer

```{r}

breast_expression <- read.csv("DataBreast/exp_array.tsv", header = T, sep = "\t")
breast_expression <- breast_expression %>% 
  select(icgc_donor_id, gene_id, normalized_expression_value)

breast_donor <- read.csv("DataBreast/donor.tsv", header = T, sep = "\t")
breast_donor <- breast_donor %>% 
  select(icgc_donor_id, donor_vital_status, disease_status_last_followup, donor_age_at_diagnosis)
 
breas_cancer <- breast_expression %>% 
  left_join(breast_donor, by = 'icgc_donor_id')

write.csv(breas_cancer, file = "AfterPreparingData/breast_expression.csv")

remove(breast_donor, breast_expression, breast_cancer)

matrix <- read.csv("AfterPreparingData/breast_expression.csv")

matrix <- matrix %>% 
  select(gene_id, normalized_expression_value, donor_vital_status, icgc_donor_id)

matrix <- aggregate(matrix$normalized_expression_value, list(matrix$gene_id, matrix$icgc_donor_id, matrix$donor_vital_status), FUN=median)

matrix <- matrix %>% 
  rename(gene_id = Group.1,
         icgc_donor_id = Group.2,
         donor_vital_status = Group.3,
         normalized_expression_value = x)

matrix <- pivot_wider(matrix, names_from = gene_id, values_from = normalized_expression_value , values_fill = NA)

write.csv(matrix, file = "AfterPreparingData/df_after_aggregate.csv")

```

#Requirements 2

```{r}

library(rpart)
library(rpart.plot)
library('doParallel')
library(caret)

```

##Load data breast


```{r}


matrix_breast <- read.csv("AfterPreparingData/df_after_aggregate.csv", row.names = 1)

matrix_breast <- matrix_breast %>% 
  select(-icgc_donor_id) 
 

matrix_breast$donor_vital_status <- as.factor(matrix_breast$donor_vital_status)


```

#Pararellation

```{r}
cl <- makeCluster(4) 
registerDoParallel(cl)
```

##Decision tree - breast cancer 

```{r}

tree_breast <- rpart(as.matrix(matrix_breast$donor_vital_status) ~ as.matrix(matrix_breast[, colnames(matrix_breast) != "donor_vital_status"]))

tiff("Figures/rpart_breast_expression.jpeg" , units="in", width=10, height=5, res=300)

t_breast <- rpart.plot(tree_breast)

dev.off()


stopCluster(cl)
```

##Training/Validation data dividing - breast

```{r}

split <- createDataPartition(y = matrix_breast$donor_vital_status, p = 0.8)
train_cancer <- matrix_breast[split$Resample1,]
valid_cancer <- matrix_breast[-split$Resample1,]
remove(matrix_breast)

data <- train_cancer
remove(train_cancer)

tree_1 <- rpart(as.matrix(donor_vital_status) ~ as.matrix(data[, colnames(data) != "donor_vital_status"]), data = data)

```
#Training_data_confusion_matrix

```{r}

pred_risk_train <- predict(object = tree_1, newdata = data, type = "class")

confusionMatrix(data = pred_risk_train, reference = data$donor_vital_status, positive = "alive", mode = "everything")
```

#Validation_data_confusion_matrix

```{r}

data <- valid_cancer

pred_risk_valid <- predict(object = tree_1, newdata = data, type = "class")

confusionMatrix(data = pred_risk_valid, reference = data$donor_vital_status,positive = "alive", mode = "everything")

remove(tree_1, split, data, valid_cancer, pred_risk_train, pred_risk_valid)
```


##Multiple linear regression

```{r}

matrix_breast <- matrix_breast %>% 
  select('UTP14A', 'STARD8','KCNV2','IREB2','MED18','IFNA10','UBE2NL', 'donor_vital_status') 

MLR_breast <- glm(as.matrix(as.numeric(matrix_breast$donor_vital_status)) ~ as.matrix(matrix_breast[, colnames(matrix_breast) != "donor_vital_status"]))

summary(MLR_breast)

```

##Violinplot - expression compare

```{r}

ggplot(data = matrix_breast) +
  geom_violin(aes(donor_vital_status, UTP14A, fill = donor_vital_status))+
  geom_boxplot(aes(donor_vital_status, UTP14A, fill = donor_vital_status), width=0.1)

ggsave(filename = "Figures/alive_decease_breast_1.jpeg", dpi = 600)


ggplot(data = matrix_breast) +
  geom_violin(aes(donor_vital_status, STARD8, fill = donor_vital_status))+
  geom_boxplot(aes(donor_vital_status, STARD8, fill = donor_vital_status), width=0.1)

ggsave(filename = "Figures/alive_decease_breast_2.jpeg", dpi = 600)


ggplot(data = matrix_breast) +
  geom_violin(aes(donor_vital_status, KCNV2, fill = donor_vital_status))+
  geom_boxplot(aes(donor_vital_status, KCNV2, fill = donor_vital_status), width=0.1)

ggsave(filename = "Figures/alive_decease_breast_3.jpeg", dpi = 600)

  
```



##Heatmap breast

```{r}

matrix_breast_1 <- data.matrix(matrix_breast[,-1])

row.names(matrix_breast_1) <- matrix_breast$donor_vital_status

remove(matrix_breast)

jpeg("Figures/rpart_breast_expression.jpeg" , units="in", width=5, height=5, res=300)

heatmap(matrix_breast_1)

dev.off()

```


###Load and preparing data for ovarian cancer

```{r}
ovarian_expression <- read.csv("DataOvary/exp_array.tsv", header = T, sep = "\t")
ovarian_expression <- ovarian_expression %>% 
  select(icgc_donor_id, gene_id, normalized_expression_value)

ovarian_donor <- read.csv("DataOvary/donor.tsv", header = T, sep = "\t")
ovarian_donor <- ovarian_donor %>% 
  select(icgc_donor_id, donor_vital_status, disease_status_last_followup, donor_age_at_diagnosis)
 
ovarian_cancer <- ovarian_expression %>% 
  left_join(ovarian_donor, by = 'icgc_donor_id')

write.csv(ovarian_cancer, file = "AfterPreparingData/ovarian_expression.csv")

remove(ovarian_donor, ovarian_expression, ovarian_cancer)

matrix <- read.csv("AfterPreparingData/ovarian_expression.csv")

matrix <- matrix %>% 
  select(gene_id, normalized_expression_value, donor_vital_status, icgc_donor_id)

matrix <- aggregate(matrix$normalized_expression_value, list(matrix$gene_id, matrix$icgc_donor_id, matrix$donor_vital_status), FUN=median)

matrix <- matrix %>% 
  rename(gene_id = Group.1,
         icgc_donor_id = Group.2,
         donor_vital_status = Group.3,
         normalized_expression_value = x)

matrix <- pivot_wider(matrix, names_from = gene_id, values_from = normalized_expression_value)

write.csv(matrix, file = "AfterPreparingData/df_after_aggregate_ovarian.csv")


```

##Load data ovary

```{r}


matrix_ovary <- read.csv("AfterPreparingData/df_after_aggregate_ovarian.csv", row.names = 1)

matrix_ovary <- matrix %>% 
  select(-icgc_donor_id) 
 

matrix_ovary$donor_vital_status <- as.factor(matrix_ovary$donor_vital_status)


```


##Decision tree - breast cancer 

```{r}

cl <- makeCluster(4) 
registerDoParallel(cl)

tree_ovary <- rpart(as.matrix(matrix_ovary$donor_vital_status) ~ as.matrix(matrix_ovary[, colnames(matrix_ovary) != "donor_vital_status"]))

jpeg("Figures/rpart_ovary_expression.jpeg" , units="in", width=10, height=5, res=300)

t_ovary <- rpart.plot(tree_ovary)

dev.off()

```


##Training/Validation data dividing - breast

```{r}

matrix_ovary <- matrix_ovary %>% 
  filter(donor_vital_status != '')

split <- createDataPartition(y = matrix_ovary$donor_vital_status, p = 0.8)
train_cancer <- matrix_ovary[split$Resample1,]
valid_cancer <- matrix_ovary[-split$Resample1,]


data <- train_cancer
remove(train_cancer)


tree_1 <- rpart(as.matrix(donor_vital_status) ~ as.matrix(data[, colnames(data) != "donor_vital_status"]), data = data)

```
#Training_data_confusion_matrix

```{r}

pred_risk_train <- predict(object = tree_1, newdata = data, type = "class")

confusionMatrix(data = pred_risk_train, reference = data$donor_vital_status, positive = "alive", mode = "everything")
```

#Validation_data_confusion_matrix

```{r}

data <- valid_cancer

pred_risk_valid <- predict(object = tree_1, newdata = data, type = "class")

confusionMatrix(data = pred_risk_valid, reference = data$donor_vital_status,positive = "alive", mode = "everything")

remove(tree_1, split, data, valid_cancer, pred_risk_train, pred_risk_valid)
```


##Multiple linear regression

```{r}

matrix_ovary <- matrix_ovary %>% 
  select('MON1A', 'SGOL2','PSRC1','MPZL2','C17orf89','DMRTA1','LOC253012','SRF', 'ACTR6','GJB1','ISG20L2','ASB16','VCAM1','IL5','CACNB1','NSF','ZNF780A',
         'KLHL10','ZC3H14','FBXW10','NR2F1','SMPD3', 'donor_vital_status') 

MLR_ovary <- glm(as.matrix(as.numeric(matrix_ovary$donor_vital_status)) ~ as.matrix(matrix_ovary[, colnames(matrix_ovary) != "donor_vital_status"]))

summary(MLR_ovary)

```


##Violinplot - expression compare

```{r}

matrix_ovary <- matrix_ovary %>% 
  filter(donor_vital_status != '')

ggplot(data = matrix_ovary) +
  geom_violin(aes(donor_vital_status, KLHL10, fill = donor_vital_status))+
  geom_boxplot(aes(donor_vital_status, KLHL10, fill = donor_vital_status), width=0.1)

ggsave(filename = "Figures/alive_decease_ovary_1.jpeg", dpi = 600)

ggplot(data = matrix_ovary) +
  geom_violin(aes(donor_vital_status, FBXW10, fill = donor_vital_status))+
  geom_boxplot(aes(donor_vital_status, FBXW10, fill = donor_vital_status), width=0.1)

ggsave(filename = "Figures/alive_decease_ovary_2.jpeg", dpi = 600)

ggplot(data = matrix_ovary) +
  geom_violin(aes(donor_vital_status, NR2F1, fill = donor_vital_status))+
  geom_boxplot(aes(donor_vital_status, NR2F1, fill = donor_vital_status), width=0.1)

ggsave(filename = "Figures/alive_decease_ovary_3.jpeg", dpi = 600)

ggplot(data = matrix_ovary) +
  geom_violin(aes(donor_vital_status, PSRC1, fill = donor_vital_status))+
  geom_boxplot(aes(donor_vital_status, PSRC1, fill = donor_vital_status), width=0.1)

ggsave(filename = "Figures/alive_decease_ovary_4.jpeg", dpi = 600)

ggplot(data = matrix_ovary) +
  geom_violin(aes(donor_vital_status, MPZL2, fill = donor_vital_status))+
  geom_boxplot(aes(donor_vital_status, MPZL2, fill = donor_vital_status), width=0.1)

ggsave(filename = "Figures/alive_decease_ovary_5.jpeg", dpi = 600)

ggplot(data = matrix_ovary) +
  geom_violin(aes(donor_vital_status, LOC253012, fill = donor_vital_status))+
  geom_boxplot(aes(donor_vital_status, LOC253012, fill = donor_vital_status), width=0.1)

ggsave(filename = "Figures/alive_decease_ovary_6.jpeg", dpi = 600)


ggplot(data = matrix_ovary) +
  geom_violin(aes(donor_vital_status, ISG20L2, fill = donor_vital_status))+
  geom_boxplot(aes(donor_vital_status, ISG20L2, fill = donor_vital_status), width=0.1)

ggsave(filename = "Figures/alive_decease_ovary_7.jpeg", dpi = 600)

  
```

##Heatmap breast

```{r}

matrix_ovary_1 <- data.matrix(matrix_ovary[,-1])

row.names(matrix_ovary_1) <- matrix_ovary$donor_vital_status

remove(matrix_ovary)

jpeg("Figures/heatmap_ovary_expression.jpeg" , units="in", width=5, height=5, res=300)

hm_breast <- heatmap(matrix_ovary_1)

dev.off()

stopCluster(cl)
```