---
title: "Proteins_analysis"
author: "Jakub Kubiś"
date: "30 08 2020"
output: html_document
---

###Analysis exrpession of genes in ovarian/breast cancer

#Requirements 1

```{r}
library(tidyverse)
Sys.setenv(LANG = "en")
set.seed(123)
```

###Load and preparing data for breast cancer

```{r}
breast_proteins <- read.csv("DataBreast/protein_expression.tsv", header = T, sep = "\t")
breast_proteins <- breast_proteins %>% 
  select(icgc_donor_id, gene_name, normalized_expression_level)

breast_donor <- read.csv("DataBreast/donor.tsv", header = T, sep = "\t")
breast_donor <- breast_donor %>% 
  select(icgc_donor_id, donor_vital_status, disease_status_last_followup, donor_age_at_diagnosis)
 
breast_cancer <- breast_proteins %>% 
  left_join(breast_donor, by = 'icgc_donor_id')

write.csv(breast_cancer, file = "AfterPreparingData/breast_protein_expression.csv")

remove(breast_donor, breast_proteins, breast_cancer)

proteins <- read.csv("AfterPreparingData/breast_protein_expression.csv")

proteins <- proteins %>% 
  select(gene_name, normalized_expression_level, donor_vital_status, icgc_donor_id)


proteins <- aggregate(proteins$normalized_expression_level, list(proteins$gene_name, proteins$icgc_donor_id, proteins$donor_vital_status), FUN=median)

proteins <- proteins %>% 
  rename(gene_name = Group.1,
         icgc_donor_id = Group.2,
         donor_vital_status = Group.3,
         normalized_expression_level = x)

proteins <- pivot_wider(proteins, names_from = gene_name, values_from = normalized_expression_level , values_fill = NA)

write.csv(proteins, file = "AfterPreparingData/breast_protein_expression_wide.csv")

```

#Requirements 2

```{r}

library(rpart)
library(rpart.plot)
library('doParallel')
library(caret)

```

##Load data breast


```{r}


proteins_breast <- read.csv("AfterPreparingData/breast_protein_expression_wide.csv", row.names = 1)

proteins_breast <- proteins_breast %>% 
  select(-icgc_donor_id) 
 

proteins_breast$donor_vital_status <- as.factor(proteins_breast$donor_vital_status)


```

#Parallelation 

```{r}

cl <- makeCluster(4) 
registerDoParallel(cl)

```


##Decision tree - breast cancer 

```{r}

tree_breast <- rpart(as.matrix(proteins_breast$donor_vital_status) ~ as.matrix(proteins_breast[, colnames(proteins_breast) != "donor_vital_status"]))

tiff("Figures/rpart_breast_proteins.jpeg" , units="in", width=10, height=5, res=300)

t_breast <- rpart.plot(tree_breast)

dev.off()


```

##Training/Validation data dividing - breast

```{r}
split <- createDataPartition(y = proteins_breast$donor_vital_status, p = 0.8)

train_cancer <- proteins_breast[split$Resample1,]
valid_cancer <- proteins_breast[-split$Resample1,]

data <- train_cancer
remove(train_cancer)

tree_1 <- rpart(as.matrix(donor_vital_status) ~ as.matrix(data[, colnames(data) != "donor_vital_status"]), data = data)
```
#Training_data_confusion_matrix

```{r}

pred_risk_train <- predict(object = tree_1, newdata = data, type = "class")

confusionMatrix(data = pred_risk_train, reference = data$donor_vital_status, positive = "alive", mode = "everything")
```

#Validation_data_confusion_matrix

```{r}

data <- valid_cancer

pred_risk_valid <- predict(object = tree_1, newdata = data, type = "class")

confusionMatrix(data = pred_risk_valid, reference = data$donor_vital_status,positive = "alive", mode = "everything")

remove(tree_1, split, data, valid_cancer, pred_risk_train, pred_risk_valid)
```


##Multiple linear regression

```{r}

proteins_breast <- proteins_breast %>% 
  select('FN1', 'CCNE1','ITGA2','BAK1','PIK3CA', 'donor_vital_status') 

MLR_breast <- glm(as.matrix(as.numeric(proteins_breast$donor_vital_status)) ~ as.matrix(proteins_breast[, colnames(proteins_breast) != "donor_vital_status"]))

summary(MLR_breast)




```

##Violinplot - protein compare

```{r}

ggplot(data = proteins_breast) +
  geom_violin(aes(donor_vital_status, FN1, fill = donor_vital_status))+
  geom_boxplot(aes(donor_vital_status, FN1, fill = donor_vital_status), width=0.1) +
  ylab("Normalized_expression") +
  xlab("Disease_status")+
  ggtitle("FN1 - protein") +
  theme(axis.text.x = element_text(angle = 50, vjust = 1, hjust=1))

ggsave(filename = "Figures/alive_decease_breast_proteins_1.jpeg", dpi = 600)

  
```



##Heatmap breast

```{r}

m <- proteins_breast %>% 
  select(-donor_vital_status)

matrix_breast_proteins <- data.matrix(m) 
  

row.names(matrix_breast_proteins) <- proteins_breast$donor_vital_status

remove(m)

tiff("Figures/rpart_breast_expression.jpeg" , units="in", width=5, height=5, res=300)

heatmap(matrix_breast_proteins)

dev.off()

stopCluster(cl)

```

